Towards practical dynamic induction control of wind farms: analysis of optimal control dynamics and sinusoidal induction control of first-row turbines

### 2018-01-14 draft.tex
Finished draft introduction

\title{Towards practical dynamic induction control of wind farms: analysis of optimal control dynamics and mimicry of first-row behavior}
\runningtitle{Towards practical dynamic induction control of wind farms}

@@

\introduction  %% \introduction[modified heading if necessary]
Wake interactions between turbines within a wind farm cause reduced power extraction and increased turbine loading in downstream rows. The current control paradigm in such farms optimizes performance at the wind-turbine level and does not account for these interactions, resulting in sub-optimal wind-farm efficiency. In contrast, control strategies at the farm level allow to influence wake interactions and promise to improve overall wind-farm performance: by redirecting propagating wakes (yaw control, see e.g. Refs YYY) or by affecting the induced wake velocity deficits (axial induction control, see e.g. Refs XXX) wind conditions for downstream turbines can be improved significantly. In recent work, Munters \& Meyers \cite{munters2017optimal} ...

Korte beschrijving van methodologie, sleutelresultaat, problemen, zie abstract...

The current paper presents efforts on understanding optimal control dynamics and identifying the flow mechanisms behind the power increase observed in the optimal induction control simulations discussed in Chapter~\ref{ch:opt_induction}, with the aim of harnessing them in tractable dynamic induction wind-farm controllers. Throughout the analysis, the focus will lie on the overinductive fast-response case C3t5, as it was observed to exhibit similar flow statistics as the maximum yield case C3t0, but with smoother thrust force dynamics.

The analysis presented in the current paper will lead to the identification of a coherent flow mechanism in the wake of optimally controlled first-row turbines, responsible for part of the power increase in the optimal control simulations. This resulted in a new derived simplified dynamic axial induction control strategy for first-row turbines producing a robust power increase in the second row, irrespective of precise local flow details. However, it will be shown that the controls responsible for the remainder of the power gains in the optimal control simulations are tuned to specific flow conditions. This complicates the analysis of downstream rows, and has thus far impeded the derivation of similar control strategies for the latter.

The outline of the paper is as follows: first, Section~\ref{sec:analysis_controls_analysis} presents an analysis of the control and thrust force dynamics, and performs some numerical experiments to elucidate characteristics of the optimal controls. It will be shown that the coherent behavior of turbines situated in the first row of the wind farm stands out from their downstream counterparts. Hereafter, Section~\ref{sec:analysis_flow_vis} identifies the shedding of vortex rings from the first row based on a flow visualization. Further, a simple sinusoidal thrust control approach is presented that successfully mimics this process with a robust increase in power extraction extraction in the second row. Next, Section~\ref{sec:analysis_intermediate} shortly discusses the behavior of the intermediate rows, i.e. turbines which have both upstream and downstream neighbors, for which similar simple control strategies have not been identified thus far. In conclusion, Section~\ref{sec:analysis_summ} summarizes the main findings of this paper.

@@

@@

\begin{acknowledgements}
The authors acknowledge funding by the European Research Council (grant no. 306471). The computational resources and services used in this work were provided by the VSC (Flemish Supercomputer Center), funded by the Research Foundation - Flanders (FWO) and the Flemish Government ? department EWI.
\end{acknowledgements}

>>>>

\title{Towards practical dynamic induction control of wind farms: analysis of optimal control dynamics and sinusoidal induction control of first-row turbines}
\runningtitle{Towards practical dynamic induction control of wind farms}

@@

\introduction  %% \introduction[modified heading if necessary]
Wake interactions between turbines within a wind farm cause reduced power extraction and increased turbine loading in downstream rows. The current control paradigm in such farms optimizes performance at the wind-turbine level and does not account for these interactions, resulting in sub-optimal wind-farm efficiency. In contrast, control strategies at the farm level allow to influence wake interactions and promise to improve overall wind-farm performance by improving wind conditions for downstream turbines, i.e. by redirecting propagating wakes (yaw control, see e.g. \citealp{fleming2014evaluating, gebraad2016wind, campagnolo2016wind}) or by affecting the induced wake velocity deficits (axial induction control, see e.g. \citealp{nilsson2015large, annoni2016analysis, bartl2016experimental}). A more exhaustive survey of wind-farm control in a broader context can be found in \cite{knudsen2015survey} and \cite{boersma2017tutorial}.

In contrast to the studies cited above, which focus on static setpoint control of wind farms, \cite{goit2015optimal} introduced a dynamic induction control approach based on large-eddy simulations and adjoint gradient optimization. The methodology was applied to the asymptotic case of a fully-developed `infinite' aligned wind farm, and power gains of about 16\% were quantified. Later, this approach was also used in induction control studies of wind farms with entrance effects in \cite{goit2016optimal} and, more recently, in \cite{munters2017optimal}, where similar gains in the order of 15\%--20\% were achieved. 
It is important to note that the computational cost of this dynamic induction control methodology is orders of magnitude too high for direct implementation, and the complex optimal control signals generated by the simulations are highly tuned to specific flow events and cannot be translated to practical flow conditions. However, the methodology allows to assess the theoretical potential for wind-farm control, and an increased understanding of the flow physics of controlled wind farms could lead to simplified control strategies that could be applied in practice.

Very recently, the methodology of \cite{goit2015optimal} was generalized to include dynamic yaw control in \cite{munters2018dynamic}. In this study, induction control and yaw control were compared for a relatively small aligned wind farm, and yaw control was found to yield higher power gains for this setup. Furthermore, the high potential of combined induction and yaw control was quantified, and analysis of the yaw control signals allowed to identify practical simplified dynamic yaw control strategies. The search for similar practical control strategies for yaw control has remained unsuccessful to date.

The current paper presents efforts on understanding optimal control dynamics and identifying the flow mechanisms behind the power increase observed in the optimal induction control simulations in \cite{munters2017optimal} (further denoted as MM17), with the aim of harnessing them in tractable dynamic induction wind-farm controllers. The analysis presented here will lead to the identification of a coherent flow mechanism in the wake of optimally controlled first-row turbines, responsible for part of the power increase in the optimal control simulations. This resulted in a new derived simplified dynamic axial induction control strategy for first-row turbines producing a robust power increase in the second row, irrespective of precise local flow details. However, it will be shown that the controls responsible for the remainder of the power gains in the optimal control simulations are tuned to specific flow conditions. This complicates the analysis of downstream rows, and has thus far impeded the derivation of similar control strategies for the latter.

The outline of the paper is as follows: first, Section~\ref{sec:optimal_control_sims} discusses the numerical setup and optimal control simulations of MM17 that will be further analysed in the current paper. Section~\ref{sec:analysis_controls_analysis} presents an analysis of the control and thrust force dynamics, and performs some numerical experiments to elucidate characteristics of the optimal controls. It will be shown that the coherent behavior of turbines situated in the first row of the wind farm stands out from their downstream counterparts. Hereafter, Section~\ref{sec:analysis_flow_vis} identifies the shedding of vortex rings from the first row based on a flow visualization. Further, a simple sinusoidal thrust control approach is presented that successfully mimics this process with a robust increase in power extraction extraction in the second row. Next, Section~\ref{sec:analysis_intermediate} shortly discusses the behavior of the intermediate rows, i.e. turbines which have both upstream and downstream neighbors, for which similar simple control strategies have not been identified thus far. In conclusion, Section~\ref{sec:analysis_summ} summarizes the main findings of this paper.

@@

Throughout the analysis, the focus will lie on the overinductive fast-response case C3t5, as it was observed to exhibit similar flow statistics as the maximum yield case C3t0, but with smoother thrust force dynamics.

@@

\begin{acknowledgements}
The authors acknowledge funding by the European Research Council (grant no. 306471). The computational resources and services used in this work were provided by the VSC (Flemish Supercomputer Center), funded by the Research Foundation - Flanders (FWO) and the Flemish Government, department EWI.
\end{acknowledgements}
